---
title: "The Amazon Bessemer Election and NLRB Election Reruns"
description: "What can historical elections tell us about the union's chances?"
summary: >
    If a new election is ordered in the RWDSU/Amazon case, it will be just as 
    bitterly contested as the first election.  But the outcome of rerun elections
    is often different than the inital election.  Here's what we might expect
    from an election rerun.
date: 2021-04-08
output: 
    md_document:
        preserve_yaml: true
slug: nlrb-reruns
images:
    - ./2021/04/nlrb-reruns/cover.png
knit: 'local({source("~/Documents/Code/website/render_rmd.R", local=(e<-new.env())); e$render_site_post})'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, fig.path="figures/", fig.retina=2,
                      dev.args=list(bg="transparent"), dev="ragg_png")

knitr::knit_hooks$set(class = function(before, options, envir) {
  if (before) sprintf("<figure class = '%s'>", options$class) else "</figure>"
})
```

<span class="leadin">The union election in Bessemer, Alabama</span>, where
Amazon warehouse workers are trying to join the Retail, Wholesale, and
Department Store Union (RWDSU), has garnered international attention.  
After a relatively long mail-in election and a ballot count prolonged by
Amazon's ballot challenges, it appears the union has lost the initial count.

```{r message=F}
library(here)
library(wacolors)
library(scales)
library(tidyverse)

d = read_rds(here("data/one_union_elec.rds"))

th = theme_minimal(base_family="Overpass", base_size=12) +
    theme(plot.title=element_text(face="bold"))
```

# Things Could Change in the Final Count

[challenged ballots]

# What Happens in Rerun Elections
The NLRB's review machinery is excruciatingly slow, and we can expect the path
to a second election to be delayed at every step by Amazon's management lawyers.
In that time, RWDSU will continue to organize, and the unit in the Bessemer
warehouse will experience significant turnover (especially given Amazon's 
abysmal turnover rates).  By the time the workers head to the polls again, 
in other words, the situation could be very different.

```{r}
d_re = group_by(d, case) %>%
    filter("Rerun" %in% tally_type) %>%
    drop_na(eligible, votes_yes, votes_no) %>%
    arrange(tally_date, .by_group=TRUE) %>%
    mutate(tally_type = str_to_lower(tally_type),
           ballot_type = if_else(ballot_type=="Revised Single Labor Org", "revised", "single"),
           votes_challenged = coalesce(votes_challenged, 0),
           void_ballots = coalesce(void_ballots, 0)) %>%
    filter(ballot_type=="single") %>%
    filter(any("initial" %in% tally_type), any("rerun" %in% tally_type)) %>%
    # get the first initial and last rerun
    filter(row_number() %in% c(which.max(tally_type=="initial"), 
                               n()+1-which.max(rev(tally_type)=="rerun"))) %>%
    select(-reason_closed, -runoff_required) %>%
    distinct() %>%
    ungroup()

d_re_sum = d_re %>%
    pivot_wider(c(case:date_filed, region, union), 
                names_from=tally_type,
                values_from=c(eligible, void_ballots, votes_yes:challenges_determ)) %>%
    transmute(state=state, region=region, 
              date_filed=date_filed, date_closed=date_closed,
              case_days = as.integer(date_closed - date_filed),
              unit_size = pmax(eligible_rerun, eligible_initial, na.rm=T),
              votes_counted_initial = coalesce(votes_counted_initial,
                                               votes_yes_initial + votes_no_initial),
              votes_counted_rerun = coalesce(votes_counted_rerun,
                                               votes_yes_rerun + votes_no_rerun),
              pct_initial = votes_yes_initial / votes_counted_initial,
              pct_rerun = votes_yes_rerun / votes_counted_rerun,
              rerun_chg = pct_rerun - pct_initial,
              chal_initial = votes_challenged_initial,
              chal_rerun = votes_challenged_rerun)

drop_na(d_re_sum, unit_size, rerun_chg) %>%
ggplot(aes(pct_initial, pct_rerun, size=unit_size, color=case_days)) +
    geom_hline(yintercept=0.5, lty="dashed") +
    geom_vline(xintercept=0.5, lty="dashed") +
    geom_smooth(aes(group=pct_initial>0.5), method=lm, formula=y~x, color="black") +
    geom_point() +
    scale_x_continuous(labels=percent) +
    scale_y_continuous(labels=percent) +
    scale_size_continuous(labels=comma)  +
    scale_color_continuous(labels=comma, trans="log10")  +
    labs(title="Anything can happen",
         subtitle="Union vote share, initial and rerun elections",
         x="Union vote in initial election", 
         y="Union vote in rerun election", 
         size="Eligible voters",
         color="Length of\ncase (days)") +
    th
```
